---
title: "Better Movie Reviews? The problem with Rotten Tomatoes, and how to fix it."
author: "Jack Cunningham"
format: html
editor: visual
echo:   false
message: false
warning: false
---

```{r, loading data}
modeling_data <- readRDS("D:/Portfolio Projects/rotten_tomatoes_empirical_bayes/Data/Processed/mixed_beta_output.RData")
posterior_stats <- readRDS("D:/Portfolio Projects/rotten_tomatoes_empirical_bayes/Data/Processed/posterior_stats.RData")
load("D:/Portfolio Projects/rotten_tomatoes_empirical_bayes/Data/Processed/movie_df.RData")
mixed_beta_output <- readRDS("D:/Portfolio Projects/rotten_tomatoes_empirical_bayes/Data/Processed/mixed_beta_output.RData")
critic_posterior_stats <- posterior_stats$critic_posterior_stats

```

```{r, loading functions}
helper_functions <- readRDS("D:/Portfolio Projects/rotten_tomatoes_empirical_bayes/Data/Functions/fit_mixed_beta_helpers.RData")
sample_from_stored_posterior <- helper_functions$sample_from_stored_posterior
```

```{r}
mix_beta_density <- helper_functions$mix_beta_density
```

```{r, loading libraries}
library(tidyverse)
library(DT)
```

# Research Motivation & Question

Let's find the the best movie based on tomatometer.

```{r}
perfect_scores <- critic_posterior_stats |> 
  filter(tomatometer_rating == 100) |> 
  select(movie_title,tomatometer_rating,tomatometer_count)
datatable(perfect_scores, colnames = c("Title","Tomatometer","Number of Critic Reviews"))
```

Hmm... there are nine hundred movies with a perfect tomatometer score, but most of them don't look familiar at all. Why is that? It's because not all movies have the same number of critic reviews. In fact, for movies with perfect reviews over fifty percent have less than ten critic reviews.

```{r}
perfect_score_lot_df <- critic_posterior_stats |> 
  mutate(perfect = factor(ifelse(tomatometer_rating == 100,"Perfect","Not Perfect"),levels = c("Perfect","Not Perfect")))
```

```{r}
perfect_score_lot_df %>%
  group_by(perfect) %>%
  mutate(w = 1 / n()) %>%   # each group sums to 1
  ungroup() %>%
  ggplot(aes(x = tomatometer_count)) +
  geom_histogram(
    aes(weight = w, y = after_stat(count)),
    binwidth = 10,
    position = "identity",
    alpha = 0.5,
    color = "white",
    fill = "#2C3E50"
  ) +
  facet_wrap(~perfect) +
  labs(x = "Number of Critic Reviews", y = "Percent within Group") +
  xlim(0,200) +
  theme_minimal()
```

With that in mind, how should we adjust our search? A natural idea is introducing a cutoff, for instance only looking at movies with over fifty critic reviews.

However with this we run into issues, what if we were interested at looking at movies that are just coming out? What about movies before the internet age that don't have enough reviews digitized and captured by rotten tomatoes? They would be ignored due to this cutoff. Is there a way to do this better?

# Empirical Bayes

A helpful technique for this situation is Empirical Bayes. As opposed to traditional Bayes, Empirical Bayes estimates the prior with the data itself. The main constraint is whether you have enough data, as this is an approximation to more exact Bayesian methods. Thankfully, with over seventeen thousand movies we certainly have enough data.

Empirical Bayes learns the distribution of movie reviews, and then will combine that information with the observed reviews from critics for each movie. This allows for us to compare movies to each other without being concerned about the number of critic reviews, and it is intuitive as we all know that a movie with a single positive critic review is not truly a perfect movie. Let's use this method.

#### Estimating a Prior

The first step is to take a look at the distribution of tomatometer for movies with a certain break point of critic reviews, this lets us filter out noise (like movies with tomato meters of zero or one hundred with only a few reviews).

```{r, filtered movies}
reduced_movies <- movie_df |> 
  mutate(p = tomato_meter_fresh_critics_count / tomatometer_count) |> 
  filter(tomatometer_count > 15)
reduced_mean <- mean(reduced_movies$tomatometer_rating/100)
reduced_median <- median(reduced_movies$tomatometer_rating/100)
```

```{r, fig 1}
reduced_movies |> 
  ggplot(aes(x = tomatometer_rating/100)) +
  geom_histogram(binwidth = .05,color = "white", fill = "#2C3E50") +
  theme_minimal() +
  labs(x = "Tomatometer", y = "Count", title = "Distribution of Tomatometer",
       subtitle = "Movies with 15 or greater reviews") +
  #geom_vline(aes(xintercept = mean(tomatometer_rating/100))) +
  annotate(geom = "segment", x = reduced_mean, xend = reduced_mean, 
           y = -100, yend = -15, arrow = arrow(length = unit(3,"mm"))) +
  annotate(geom = "text", x = reduced_mean, y = -150, label = "Mean") +
  annotate(geom = "segment", x = .9, xend = .9, 
           y = -100, yend = -15, arrow = arrow(length = unit(3,"mm"))) +
  annotate(geom = "text", x = .9, y = -150, label = "Mode") +
  annotate(geom = "segment", x = reduced_median, xend = reduced_median, 
           y = -200, yend = -15, arrow = arrow(length = unit(3,"mm"))) +
  annotate(geom = "text", x = reduced_median, y = -250, label = "Median")

```

The distribution of tomatometer is left skewed, with a peak near 90. The mean is slightly left of the median, with an average score of 62 and a median score of 69. There are very few tomatometer ratings in the band of 0 - 0.05.

We fit a distribution to this observed data to form an empirical prior, in this case a mixed beta distribution will fit nicely.

```{r}
reduced_movies |> 
  ggplot(aes(x = tomatometer_rating/100)) +
  geom_histogram(aes(y = after_stat(density)),bins = 20,color = "white", fill = "#2C3E50") +
  stat_function(fun = mix_beta_density, 
              args = list(pi =  modeling_data$mm_critic$pi, alpha = modeling_data$mm_critic$alpha, beta = modeling_data$mm_critic$beta),
              color = "red") +
  labs(x = "Tomatometer", y = "", title = "Empirical Prior",
       subtitle = "Mixed Beta fit") +
  theme_minimal()
```

Now with this prior we can analyze our posterior estimates of tomatometer for each movie.

# Posterior Analysis; How have our ratings changed?

One of the most helpful aspects of Emprical Bayes is how it "shrinks" estimates, movies are pulled toward what the prior distribution would suggest. This shrinking occurs strongly for movies with only a few reviews, and lightly for movies with many reviews (as the observed data dominates the information provided by the prior). Let's see how shrinking affects the posterior estimates as the count of reviews change.

```{r}
shrinkage_df <- critic_posterior_stats |> 
  filter(tomatometer_rating < 15 | tomatometer_rating > 85) |> 
  mutate(high_scoring = ifelse(tomatometer_rating > 85, "High Score",
                               "Low Score")) |> 
  mutate(high_scoring = factor(high_scoring, c("Low Score","High Score")))
```

```{r}
shrinkage_df |> 
  ggplot(aes(x = tomatometer_rating, y = post_mean*100, col = log(tomatometer_count))) +
  geom_point() +
  geom_abline(col = "#26F7FD") +
  facet_wrap(~ high_scoring, scales = "free") +
  theme_minimal() +
  labs(x = "Observed Tomatometer Rating", y = "Adjusted Tomatometer Rating",
       title = "How Bayesian Shrinking Affects Extreme Observed Ratings",
       subtitle = "Line is the observed tomatometer") +
  scale_color_gradient(low = "#08306b", high = "#26F7FD",
                       name = "Critic Reviews",
                        breaks = log(c(10,100,500)),
                        labels = c("10","100","500")) +
  guides(color = guide_colorbar(
  barheight = 8,
  barwidth = 1
))
```

The shrinking has its greatest effect on extreme values with only a few reviews.

For instance, a movie with a tomatometer rating of zero with a single review gets pulled up to an adjusted tomatometer rating of 20. As we observed earlier, scores very close to zero happen rarely.

On the other hand, movies with a perfect tomatometer rating get pulled down when there are only a few reviews.

Now we can look at adjusted tomatometer ratings to find the movies that are actually most beloved by critics, and see their credible intervals. See the top one thousand movies according to the adjusted ratings below.

```{r}
adjusted_scores <- critic_posterior_stats |> 
  select(movie_title,post_mean,lo_95,hi_95,tomatometer_rating,tomatometer_count) |>
  arrange(desc(post_mean)) |> 
  mutate(across(c(post_mean,lo_95,hi_95),round,3)) |> 
  head(1000)

datatable(adjusted_scores, colnames = c("Title","Adjusted Tomatometer Rating","2.5%","97.5%","Original Tomatometer Rating","Number of Critic Reviews"))
```

Our adjusted rating does a much better job of showing the best movies according to critics, all of these top movies were nominated for awards!

# Comparisons

Empirical Bayes also gives us the tools for interesting tasks, like comparing movies to each other to find the best movie of a bunch.

Let's say we are in the mood for a Christmas comedy and have a list of four potential movies, but only the time to watch one. Here are our four movies:

```{r}
critics_out <- mixed_beta_output$critics_out

christmas_comedies <- critics_out |> 
  filter(movie_title %in% c("National Lampoon's Christmas Vacation",
                            "Elf","A Christmas Story","Home Alone"))

christmas_comedies_output <- critic_posterior_stats |>
  filter(movie_title %in% c("National Lampoon's Christmas Vacation",
                            "Elf","A Christmas Story","Home Alone")) |> 
  select(movie_title,post_mean,lo_95,hi_95,tomatometer_rating,tomatometer_count) |>
  mutate(across(c(post_mean,lo_95,hi_95),round,3))

datatable(christmas_comedies_output, colnames = c("Title","Adjusted Tomatometer Rating","2.5%","97.5%","Original Tomatometer Rating","Number of Critic Reviews"),
          options = list(searching = FALSE, lengthChange = FALSE,
                         paging = FALSE, info = FALSE))
```

We could just select the one with the highest adjusted tomatometer rating, but finding the probability that each movie is the best of the bunch can give a good idea of how far apart each movie really is.

```{r}
set.seed(10)
ndraw <- 100000
df_draws <- christmas_comedies |> 
  mutate(
     draws = pmap(
      list(w1 = w1, a1 = a1, b1 = b1, a2 = a2, b2 = b2),
      ~sample_from_stored_posterior(w1 = ..1, a1 = ..2, b1 = ..3, 
                                    a2 = ..4, b2 = ..5, S = ndraw)
     )
 )
```

```{r}
draws_long <- df_draws |> 
  select(movie_title, draws) |> 
  unnest_longer(draws, values_to = "theta") |> 
  group_by(movie_title) |> 
  mutate(draw = row_number()) |> 
  ungroup()
```

```{r}
draws_long |> 
  ggplot(aes(x = theta, color = movie_title)) +
  geom_density() +
  labs(title = "Posterior Distributions", x = "Adjusted Tomatometer Rating",
       y = "Density") +
  theme_minimal()
```

```{r}
best_prob <- draws_long |> 
  group_by(draw) |> 
  mutate(max_draw = max(theta)) |> 
  ungroup() |> 
  mutate(best = theta == max_draw) |> 
  group_by(movie_title) |> 
  summarise(best_probability = mean(best))
datatable(best_prob,colnames = c("Title","Probability of being best"),
          options = list(searching = FALSE, lengthChange = FALSE,
                         paging = FALSE, info = FALSE))
```

It seems that A Christmas Story has the highest probability of being the best movie according to critics, but Elf has a probability of about seventeen percent. These two movies are in contention for being the best Christmas comedy movie according to critics.

With this output we could setup a rule. For instance we could watch the trailers of the movies with greater than a ten percent chance of being the best movie according to critics before making a decision on what to watch.
